Location and Scale-Invariant Power Transformations for Transforming Data to

### 2025-03-31 manuscript_rmarkdown/manuscript.Rmd
Inserted new values.

To determine the weighting function parameters for each of the nine combinations, $m_d=100$ asymmetric generalised normal distributions were drawn.
Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
$n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, with $\gamma \sim U\left(2, 4\right)$, i.e., between $100$ and $10000$ values are drawn to create $\mathbf{X}_i$.

Outlier values were then drawn to randomly replace a fraction of the values of $\mathbf{X}_i$.
This was repeated $m_{\text{out}} = 10$ times, with outlier fractions regularly spaced in $[0.00, 0.10]$.
Thus up to 10 percent of the values could be replaced by outliers.
Outlier values were set according to @Tukey1977-xm, as follows.
Let $x^{*} \sim U\left(-2, 2\right)$.
Then the corresponding outlier value was:

```{=latex}
\begin{equation}
x_{out} =
\begin{cases}
Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
\end{cases}
\end{equation}
```

$Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
Outlier values randomly replaced values in $\mathbf{X}_i$ to create $\mathbf{X}_{i,j}$, with $j \in \left\{ 1, \ldots, m_{\text{out}} \right\}$

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for robust transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:

```{=latex}
\begin{equation}
\label{eqn:minimisation-weighting}
\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{\text{out}}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
\end{equation}
```

@@

```{=latex}
\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations. $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. The {optimal} column shows the optimal parameter values. The \textit{loss} column shows the loss achieved by each method, under optimised parameters. Lower loss indicates better robustness against outliers.}
\label{tab:optimal-weighting-parameters-box-cox}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 771 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.86 & ---  & ---       & ---  & 561 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.83 & 0.95 & $(0, 1]$  & 0.92 & 564 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.76 & 0.95 & $(0, 1]$  & 0.95 & 560 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 1.12 & ---  & ---       & ---  & 1153 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.38 & 1.96 & $(0, 10]$ & 4.48 & 1178 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.21 & 1.96 & $(0, 10]$ & 4.92 & 1135 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 2.18 & ---  & ---       & ---  & 1839 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.29 & 1.00 & $(0, 10]$ & 1.47 & 1764 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.38 & 1.00 & $(0, 10]$ & 1.41 & 1583 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{=latex}
\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations. $p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values, and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. The {optimal} column shows the optimal parameter values. The \textit{loss} column shows the loss achieved by each method, under optimised parameters. Lower loss indicates better robustness against outliers.}
\label{tab:optimal-weighting-parameters-yeo-johnson}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 364 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.95 & ---  & ---       & ---  & 224 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.88 & 0.95 & $(0, 1]$  & 0.97 & 212 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.94 & 0.95 & $(0, 1]$  & 0.95 & 218 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 1.09 & ---  & ---       & ---  & 392 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.24 & 1.96 & $(0, 10]$ & 4.62 & 396 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.15 & 1.96 & $(0, 10]$ & 6.05 & 413 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 1.63 & ---  & ---       & ---  & 746 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.50 & 1.00 & $(0, 10]$ & 1.79 & 731 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.51 & 1.00 & $(0, 10]$ & 1.57 & 612 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

>>>>

To determine the weighting function parameters for each of the nine combinations,
$m_d=100$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(2, 4\right)$, resulting in a set of sequences with between $100$ and $10000$ elements.

Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
Outlier values were set according to @Tukey1977-xm, as follows.
Let $x^{*} \sim U\left(-2, 2\right)$.
Then the corresponding outlier value was:

\begin{equation}
x_{out} =
\begin{cases}
Q_1 - \left(1.5 - x^{*} \right) \text{IQR} & \text{if } x^{*} < 0 \\
Q_3 + \left(1.5 + x^{*} \right) \text{IQR} & \text{if } x^{*} \geq 0
\end{cases}
\end{equation}

$Q_1$, $Q_3$ and $\text{IQR}$ are the first quartile, third quartile and interquartile range of $\mathbf{X}_i$, respectively.
Outlier values randomly replaced elements in $\mathbf{X}_i$.

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised $\sum_{i=1}^{m_d = 100}\tau_{n, \kappa = 0.70, i}$, i.e. the mean of residual 
errors of the central 70% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.

@@

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. 
The \textit{optimal} column shows the optimal parameter values.
The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
Lower loss indicates better transformation to central normality in the presence of outliers.}
\label{tab:optimal-weighting-parameters-box-cox}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 4.98 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.75 & ---  & ---       & ---  & 2.83 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.82 & 2.63 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.89 & 2.63 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 2.33 & ---  & ---       & ---  & 5.04 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.20 & 1.96 & $(0, 10]$ & 4.90 & 4.96 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.22 & 1.96 & $(0, 10]$ & 5.12 & 5.23 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 0.75 & ---  & ---       & ---  & 5.07 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.83 & 1.00 & $(0, 10]$ & 0.84 & 5.18 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.75 & 1.00 & $(0, 10]$ & 0.76 & 5.11 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
The \textit{optimal} column shows the optimal parameter values. 
The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
Lower loss indicates better transformation to central normality in the presence of outliers.}
\label{tab:optimal-weighting-parameters-yeo-johnson}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 5.09 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 2.85 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.83 & 2.46 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.04 & 0.95 & $(0, 1]$  & 0.88 & 2.45 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 3.32 & ---  & ---       & ---  & 4.50 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.42 & 1.96 & $(0, 10]$ & 4.35 & 4.08 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.53 & 1.96 & $(0, 10]$ & 5.08 & 4.18 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 1.05 & ---  & ---       & ---  & 5.79 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.04 & 1.00 & $(0, 10]$ & 1.09 & 5.89 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.07 & 1.00 & $(0, 10]$ & 1.08 & 5.81 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
